---
title: "SongPopularityPrediction"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: lumen
    highlight: espresso
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,  #suppress messages
                      warning = FALSE,  #suppress warnings
                      error = FALSE)    #suppress errors
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Load required libraries
</h3>
```{r}
library(dplyr) #dataframe
library(tidyr) #tidy data
library(ggplot2) #data visualization
library(reshape) #data restructuring and aggregation
library(reshape2) #data transformation
library(rpart) #build classificaiton and regression trees
library(rpart.plot) #scale trees at best fit
library(caret) #train different algorithms
library(tidyverse) #data presentation
library(skimr) #provide summmary statistics
library(xgboost) #xg boost modelling
library(caTools) #split train and test sets
library(Metrics) #supervised machine learning metrics
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Load Dataset
</h3>

```{r}
data<-read.csv("spotify.csv")
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The dataset was stored and called from csv file.
</h4>



<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Dataset info
</h3>
```{r}
dim(data) #dataset size
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The dataset contains 114000 rows and 21 columns.
</h4>


```{r}
glimpse(data)  #view data types
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
Checking out the data types of attributes in the dataset.
</h4>

```{r}
head(data)
#tail(data)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
Viewing the head of the dataset.
</h4>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
View column names
</h3>
```{r}
ls(data)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
Viewing the list of columns.
</h4>


<span style="
font-size:15px;
text-align:Left">
All the column names with descriptions are shown below.

* ```track_id```: The Spotify ID for the track

* ```artists```: The artists' names who performed the track. If there is more than one artist, they are separated by a ;

* ```album_name```: The album name in which the track appears

* ```track_name```: Name of the track

* ```popularity```: The popularity of a track is a value between 0 and 100, with 100 being the most popular. The popularity is calculated by algorithm and is based, in the most part, on the total number of plays the track has had and how recent those plays are. 

* ```duration_ms```: The track length in milliseconds

* ```explicit```: Whether or not the track has explicit lyrics (true = yes it does; false = no it does not OR unknown)

* ```danceability```: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable.

* ```energy```: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. 

* ```key```: The key the track is in. Integers map to pitches using standard Pitch Class notation. E.g. 0 = C, 1 = C♯/D♭, 2 = D, and so on. If no key was detected, the value is -1.

* ```loudness```: The overall loudness of a track in decibels (dB)

* ```mode```: Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.

* ```speechiness```: Speechiness detects the presence of spoken words in a track. 

* ```acousticness```: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 

* ```instrumentalness```: Predicts whether a track contains no vocals. "Ooh" and "aah" sounds are treated as instrumental in this context. Rap or spoken word tracks are clearly "vocal". The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content.

* ```liveness```: Detects the presence of an audience in the recording. Higher liveness values represent an increased probability that the track was performed live. A value above 0.8 provides strong likelihood that the track is live.

* ```valence```: A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. 

* ```tempo```: The overall estimated tempo of a track in beats per minute (BPM). 

* ```time_signature```: The time signature (meter) is a notational convention to specify how many beats are in each bar (or measure). 

* ```track_genre```: The genre in which the track belongs.
</span>


<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
DATA CLEANING
</h2>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Missing value analysis - check which columns have missing values
</h3>
```{r}
is.null(data)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
No missing values present in the dataset.
</h4>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Checking for Null values in dataset
</h3>
```{r}
sum(is.null(data))
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
No null values present in the dataset.
</h4>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Check which columns have NA values
</h3>
```{r}
sum(is.na(data))
colSums(is.na(data))
names(which(colSums(is.na(data))>0))
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
No NA values present in the dataset.
</h4>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Check for duplicated entry
</h3>
```{r}
data[duplicated(data), ] #check for duplicated rows
sum(data[duplicated(data), ])
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
No duplicated values present in the dataset.
</h4>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Missing value correlation analysis
</h3>
```{r}
data1<- data %>% select(popularity, mode, instrumentalness)
cor(data1, data1,  method = "pearson", use = "complete.obs")
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Visualization of Missing value correlation analysis
</h3>
```{r}
heatmap(cor(data1, data1,  method = "pearson", use = "complete.obs"))
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
It can be seen that these missing values do not show up significantly in the overall data set and have no strong correlation. <br>

Hence, missing value rows can be deleted directly.
</h4>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Delete those rows with missing values
</h3>
```{r}
data<- data %>% drop_na()
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
View dataset info after cleaning
</h3>
```{r}
dim(data)
tail(data)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The dataset contains 113991 rows and 21 columns after removing NA values rows.
</h4>



```{r}
colSums(is.na(data))
sum(is.na(data))
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The dataset is cleaned where no columns present with NA values. The dataset shows complete cases.
</h4>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Transform explicit variable from categorical data into binary data
</h3>
```{r}
data$explicitUpdated<-ifelse(data$explicit =="FALSE", "0", "1")
data$explicitUpdated <- as.factor(data$explicitUpdated)
class(data$explicitUpdated)
levels(data$explicitUpdated)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The ```explicit``` attribute will be changed to 0 when it is ```FALSE```, while 1 is ```TRUE```
</h4>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Transform track_genre to 5 parent_genre
</h3>
```{r}
data<-data %>%
  mutate(
    parent_genre= case_when(
      track_genre =="anime" | track_genre =="brazil" | track_genre =="british" | track_genre =="cantopop" | track_genre =="j-pop" | track_genre =="k-pop" | track_genre =="indie-pop" | track_genre =="malay" | track_genre =="mandopop" | track_genre =="power-pop" | track_genre =="pop-film" | track_genre =="pop" | track_genre =="synth-pop" | track_genre =="trip-hop" | track_genre =="j-idol" | track_genre =="dance" | track_genre =="salsa" | track_genre =="samba" | track_genre =="dancehall" | track_genre =="forro" | track_genre =="garage" | track_genre =="j-dance" | track_genre =="party" | track_genre =="tango" | track_genre =="happy" | track_genre =="disco" | track_genre =="afrobeat" | track_genre =="comedy" | track_genre =="hip-hop" | track_genre =="latin" | track_genre =="latino" | track_genre =="world-music" | track_genre =="mpb" ~ "Pop",
      track_genre =="blues" | track_genre =="dub" | track_genre =="dubstep" | track_genre =="r-n-b" | track_genre =="reggae" | track_genre =="reggaeton" | track_genre =="soul" | track_genre =="ska" ~ "R & B",
      track_genre =="alt-rock" | track_genre =="alternative" | track_genre =="emo" | track_genre =="funk" | track_genre =="hard-rock"| track_genre =="psych-rock"| track_genre =="punk-rock"| track_genre =="rock-n-roll" | track_genre =="rock" | track_genre =="rockabilly" | track_genre =="j-rock" | track_genre =="goth" | track_genre =="groove" | track_genre =="grunge" | track_genre =="indie" | track_genre =="iranian" | track_genre =="new-age" | track_genre =="black-metal" | track_genre =="death-metal" | track_genre =="heavy-metal" | track_genre =="metal" | track_genre =="metalcore" | track_genre =="grindcore" | track_genre =="punk" | track_genre =="hardcore" | track_genre =="hardstyle" | track_genre =="industrial" ~ "Rock",
      track_genre =="edm" | track_genre =="electro" | track_genre =="electronic" | track_genre =="chicago-house" | track_genre =="breakbeat" | track_genre =="club" | track_genre =="deep-house" | track_genre =="detroit-techno" | track_genre =="minimal-techno" | track_genre =="techno" | track_genre =="progressive-house" | track_genre =="trance" | track_genre =="drum-and-bass" | track_genre =="house" | track_genre =="idm" ~ "EDM",
      track_genre =="acoustic" | track_genre =="folk" | track_genre =="french" | track_genre =="german" | track_genre =="guitar" | track_genre =="sertanejo" | track_genre =="study" | track_genre =="sad" | track_genre =="singer-songwriter" | track_genre =="songwriter" | track_genre =="swedish" | track_genre =="turkish" | track_genre =="gospel" | track_genre =="children" | track_genre =="ambient" | track_genre =="chill" | track_genre =="classical" | track_genre =="disney" | track_genre =="kids" | track_genre =="opera" | track_genre =="jazz" | track_genre =="piano" | track_genre =="romance" | track_genre =="indian" | track_genre =="show-tunes" | track_genre =="sleep" | track_genre =="spanish" | track_genre =="bluegrass" | track_genre == "country" | track_genre =="honky-tonk" | track_genre =="pagode" ~ "Classical"
            )
    )
    
colSums(is.na(data))

data$parent_genre<- as.factor(data$parent_genre)
data$parent_genre_class<- as.numeric(data$parent_genre)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The ```track_genre``` attribute will be changed to ```parent_genre``` of 5 categories  (pop, r&b, rock, edm, classical) and factor data into numerical data of 1 to 5.
</h4>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Convert all attributes into numerical data
</h3>
```{r}
data$popularity<- as.numeric(data$popularity)
data$duration_ms<- as.numeric(data$duration_ms)
data$explicitUpdated<- as.numeric(data$explicitUpdated)
data$danceability<- as.numeric(data$danceability)
data$energy<- as.numeric(data$energy)
data$key<- as.numeric(data$key)
data$loudness<- as.numeric(data$loudness)
data$mode<- as.numeric(data$mode)
data$speechiness<- as.numeric(data$speechiness)
data$acousticness<- as.numeric(data$acousticness)
data$instrumentalness<- as.numeric(data$instrumentalness)
data$liveness<- as.numeric(data$liveness)
data$valence<- as.numeric(data$valence)
data$tempo<- as.numeric(data$tempo)

```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Transform '''popularity''' attribute to 4 categories
</h3>
```{r}
data$popularity_class<-ifelse(data$popularity < 25, "Not Popular",
                              ifelse(data$popularity >= 25 & data$popularity <= 50, "Less Popular",
                                     ifelse(data$popularity >= 50 & data$popularity <= 75, "Popular", "Most Popular")))

# data$popularity_class<-ifelse(data$popularity <25, "Not Popular", 
#                               ifelse(data$popularity >= 25 & data$popularity <= 50, "Popular", "Most Popular"))

class(data$popularity_class)
data$popularity_class <- as.factor(data$popularity_class)
class(data$popularity_class)
levels(data$popularity_class)
data$popularity_class<- factor((data$popularity_class), levels = c("Not Popular", "Less Popular", "Popular", "Most Popular"))
levels(data$popularity_class)

data$popularity_class<- as.numeric(data$popularity_class)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The ```popularity``` attribute will be changed to '''popularity_class''' of 4 categories (Not Popular, Less Popular, Popular, Most Popular) and factor data into numerical data of 1 to 4.
</h4>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Display summary of dataset
</h3>
```{r}
str(data)
summary(data)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
Here it consists of statistical summary of the dataset.
</h4>

<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
DATA NORMALIZATION
</h2>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Combine selected attributes needed for normalization and stored into data_trans
</h3>
```{r}
data_trans <- subset(data, select = c(key, loudness, speechiness, instrumentalness, liveness, tempo, duration_ms))
head(data_trans)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
Few attributes were selected and put into subset ```data_trans``` as it requires data normalization.
</h4>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Normalize the data types using Min-Max scaler, combine and form a new dataframe for cleaned dataset
</h3>
```{r}
data_norm <- as.data.frame(apply(data_trans, 2, function(x) (x-min(x))/(max(x)-min(x))))

data_norm$danceability <- data$danceability
data_norm$energy <- data$energy
data_norm$acousticness <- data$acousticness
data_norm$valence <- data$valence
data_norm$time_signature <- data$time_signature
data_norm$parent_genre_class <- data$parent_genre_class
data_norm$popularity_class <- data$popularity_class
head(data_norm)
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The ```key```, ```loudness```, ```speechiness```, ```instrumentalness```, ```liveness```, ```tempo```, ```duration_ms``` attributes undergoes data normalization to better organize and simplify the data and were stored into a new dataframe ```data_norm```. Other attributes that are not required to undergo data normalization were selected and copied into the new dataframe as well.
</h4>


<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
EXPLORATORY DATA ANALYSIS (EDA)
</h2>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
EDA is a critical process of investigating and exploring further insights on the dataset selected
</h3>
```{r}
ggplot(data, aes(x = popularity_class)) + geom_histogram()
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The above histogram visualized the spread of popularity values across the dataset that was separated into 4 categories.
</h4>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Generate plots for all attributes in ```data_norm```
</h3>
```{r}
for (col in 2:ncol(data_norm)) {
    hist(data_norm[,col], main=colnames(data_norm)[col],  xlab=colnames(data_norm)[col])
}
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Generate correlation analysis and plot heatmap using ```data_norm```
</h3>
```{r}
ht_cor <- cor(data_norm)
heatmap(ht_cor, scale="column", col = terrain.colors(256))
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
Here it shows the correlation among the attributes in data_norm.
</h4>



<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Relation between danceability and popularity
</h3>
```{r}
ggplot(data, aes(x = danceability, y = popularity)) +
   geom_point(alpha = 0.1) +
   geom_density_2d()
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
From the scatter plot diagram, it can be deduced that when there is higher score in danceability, it resulted in higher popularity scores.
</h4>



<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Relation between energy and popularity
</h3>
```{r}
par(mfrow = c(3, 2))

# Draw the first subgraph
plot(data$energy[1:1000], data$popularity[1:1000])

# Draw the second subgraph
plot(data$energy[301:600], data$popularity[301:600])
```

```{r}
ggplot(data, aes(x = energy, y = popularity)) +
   geom_point(alpha = 0.1) +
   geom_density_2d()
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
From the result, there is negative correlation between energy and popularity.
</h4>



<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
DATA MODELLING - EXTREME GRADIENT BOOSTING (XG BOOST)
</h2>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
XG Boost supervised machine learning prediction model; split into 70% training and 30% testing set; maximum depth for decision tree = 6; 500 maximum iterations.
</h3>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
This is another method to display the dataframe details but of one row per variable analysis.
</h3>
```{r}
skim_to_wide(data_norm)
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
* Split the normalized data into 70% training and 30% testing set using ```createDataPartition```<br>
* Define predictor and target variables for both training and testing, ```train_x```,  ```train_y```, ```test_x```, and ```test_y```<br>
* Finalized XG Boost training and testing sets for 500 iterations, ```xgb_train``` and ```xgb_test```
</h3>
```{r}
index <- createDataPartition(y = data_norm$popularity_class, 
                             p = 0.7, 
                             list = FALSE)
train <- data_norm[index, ]
test <- data_norm[-index, ]
rm(index)

#training predictor and target variables
train_x <- data.matrix(train[, -14])
train_y <- train[, 14]

#testing predictor and target variables
test_x <- data.matrix(test[, -14])
test_y <- test[, 14]

#finalized training and testing sets
xgb_train <- xgb.DMatrix(data = train_x, label = train_y)
xgb_test <- xgb.DMatrix(data = test_x, label = test_y)

```


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Store and visualize the 500 iterations into a list using the ```watchlist``` parameter 
</h3>
```{r}
#define watchlist
list_created <- list(train = xgb_train, test = xgb_test)

#show training and testing data for xg boost
model <- xgb.train(data = xgb_train, max.depth = 6, watchlist = list_created, nrounds = 500)
```

```{r}
#finalized parameters for xg boost
final <- xgboost(data = xgb_train, max.depth = 6, nrounds = 500, verbose = 0)

#make predictions from training data on testing data
pred_y <- predict(final, xgb_test)
```

<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
MODEL PERFORMANCE EVALUATION - EXTREME GRADIENT BOOSTING (XG BOOST)
</h2>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
To carry out the prediction accuracy using Mean Squared Error ```MSE```, Mean Average Error ```MAE```, and Root Mean Squared Error ```RMSE``` metrics with ```caret```
</h3>
```{r}
#evaluate model
mse = mean((test_y - pred_y)^2)
mae = caret::MAE(test_y, pred_y)
rmse = caret::RMSE(test_y, pred_y)

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)
```









<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
DATA MODELLING - MULTIPLE LINEAR REGRESSION
</h2>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Linear regression
</h3>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Remove target variable from normalized data and create a boxplot
</h3>
```{r}
temp_data <- subset(data_norm, select = -c(popularity_class))
melt_data <- melt(temp_data)

boxplot(data = melt_data, value ~ variable)
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
* Split the normalized data into 70% training and 30% testing set using ```sample.split``` from ```caTools``` <br>
* Fit the linear model using ```lm()``` function and create 
summary of the model
</h3>
```{r}
set.seed(42)

#split training and testing data
partition <- sample.split(Y = data_norm$popularity_class, SplitRatio = 0.7)
train_set <- subset(x = data_norm, partition == TRUE)
test_set <- subset(x = data_norm, partition == FALSE)

#fit linear model and summarize results
model5 <- lm(popularity_class ~ ., data = train_set)
summary(model5)
```

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Define and visualized residuals with ```ggplot``` (histogram)
</h3>
```{r}
#get residuals 
lm_residuals <- as.data.frame(residuals(model5))

#visualized residuals
ggplot(lm_residuals, aes(residuals(model5))) + geom_histogram(fill = "pink", color = "black") + theme_classic() + labs(title = "Residuals Plot")
```
<h4 style="
font-size:15px;
text-align:Right;
font-style: italic;
color: orange"> 
The obtained residuals is skewed towards the left which suggests possible errors on the right.
</h4>



<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
* Create predictions by applying the model on testing set using ```predict()``` <br>
* Store and print results into a dataframe named ```evaluation```
</h3>
```{r}
#make prediction
prediction <- predict(model5, test_set)

#convert the results into a dataframe 
evaluation <- cbind(test_set$popularity_class, prediction)
colnames(evaluation) <- c("Actual", "Predicted")
evaluation <- as.data.frame(evaluation)
head(evaluation)
```

<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
MODEL PERFORMANCE EVALUATION - MULTIPLE LINEAR REGRESSION 
</h2>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
To carry out the prediction accuracy using Mean Squared Error ```MSE```, Mean Average Error ```MAE```, and Root Mean Squared Error ```RMSE``` metrics with ```Metrics```
</h3>
```{r}
#library(Metrics)
#evaluate model
mse <- mean((evaluation$Actual - evaluation$Predicted)^2)
mae <- mae(evaluation$Actual, predict(model5))
rmse <- sqrt(mse)

cat("MSE: ", mse, "MAE: ", mae, " RMSE: ", rmse)
```










<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
DATA MODELLING - DECISION TREE 
</h2>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Split the normalized data into 70% training and 30% testing set using ```createDataPartition```
</h3>
```{r}
set.seed(10000)
#split training and testing data
index <- createDataPartition(y = data_norm$popularity_class,
                             p = 0.7,
                             list = FALSE)
train_data <- data_norm[index, ]
test_data <- data_norm[-index, ]
rm(index)

#head(train_data)
#head(test_data)
```


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Create classification model with 10 fold cross-validation for ```xval = 10``` and plot the decision tree ```plot()```
</h3>
```{r}
#classification
set.seed(100)
full_classification <- rpart(formula = popularity_class~.,
                             data = train_data,
                             method = "class",
                             xval = 10)
rpart.plot(full_classification)
```


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Display CP table for fitted decision tree with ```printcp```
</h3>
```{r}
#show summary
printcp(full_classification)
```


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Display graph of cp values to cross validated error summary with ```plotcp()```
</h3>
```{r}
#show graph
plotcp(full_classification)
```


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Prune using ```prune()``` and plot decision tree by specifying the cp (associated cost-complexity)
</h3>
```{r}
#pruning the model
data_classification <- prune(full_classification,
                             cp = full_classification$cptable[which.min(full_classification$cptable[, "xerror"]), "CP"])
rm(full_classification)

#plot results
rpart.plot(data_classification, yesno = TRUE)
```


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Create predictions by applying the model on testing set using ```predict()```
</h3>
```{r}
#make prediction
data_classification_pred <- predict(data_classification,
                                    test_data,
                                    type = "class")
plot(test_data$popularity_class, data_classification_pred, main = "Classification - Predict vs Actual", xlab = "Actual", ylab = "Predicted")
```

<h2 style="
font-size:25px;
text-align:center;
text-decoration: underline;
font-weight: bold;
color: red"> 
MODEL PERFORMANCE EVALUATION - DECISION TREE  
</h2>

<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Decision tree supervised machine learning algorithm
</h3>


<h3 style="
font-size:15px;
text-align:Left;
font-style: italic;
color: blue"> 
Create confusion matrix to evaluate the model accuracy with ```confusionMatrix```
</h3>
```{r}
#evaluate model
data_classification_confusion <- confusionMatrix(data = as.factor(data_classification_pred),
                                                 reference = as.factor(test_data$popularity_class))

data_classification_confusion
```